---
title: "Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis."
author:
  - >
    Alexey N. Shiklomanov$^1$
    Elizabeth M. Cowdery$^1$,
    Michael Bahn$^2$,
    Chaeho Byun$^3$,
    Joseph Craine$^4$,
    Andrés Gonzalez-Melo$^5$,
    Steven Jansen$^6$,
    Nathan Kraft$^7$,
    Koen Kramer$^{8}$,
    Vanessa Minden$^{9,10}$,
    Ülo Niinemets$^{11}$,
    Yusuke Onoda$^{12}$,
    Enio Egon Sosinski$^{13}$,
    Nadejda A. Soudzilovskaia$^{14}$,
    Michael C. Dietze$^1$
bibliography:
  - library.bib
  - extra.bib
  - tryrefs.bib
csl: new-phytologist.csl
output:
  bookdown::word_document2:
    toc: no
    reference_docx: ~/Projects/misc/pandoc-docx-template/custom.docx
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: yes
    toc: no
    includes:
      in_header: preamble.tex
---
<div custom-style="Affiliation">
$^1$Department of Earth & Environment, Boston University, Boston, MA 02215, USA;
$^2$Institute of Ecology, University of Innsbruck, 6020 Innsbruck, Austria;
$^3$School of Biological Sciences, Seoul National University, Gwanak-ro 1, 08826, Seoul, Seoul, South Korea;
$^4$Jonah Ventures, Manhattan, KS 66502, USA;
$^5$Universidad del Rosario. Facultad de Ciencias Naturales y Matemáticas. Carrera 26 #63B-48. Bogotá, Colombia;
$^6$Institute of Systematic Botany and Ecology, Ulm University, Ulm, Germany;
$^7$Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA;
$^{8}$Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands;
$^{9}$Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
$^{10}$Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels;
$^{11}$Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia;
$^{12}$Graduate School of Agriculture, Kyoto University, Kyoto, 606-8502, Japan;
$^{13}$Embrapa Temperate Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
$^{14}$Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands
</div>

<div custom-style="Single space">
Corresponding author:

Alexey Shiklomanov,

telephone: +1 (302) 897-3348,

email: ashiklom\@bu.edu

Key words: Functional trade-off; hierarchical modeling; trait variation; ecological modeling; leaf morphology; leaf biochemistry
</div>

Section                   Word count
--------                  --------------      ------------
Introduction              1060                Figures: 5 (all colored)
Materials and Methods     1695                Tables: 2
Results                   1108                Supplementary figures: 2
Discussion                1907                Supplementary tables: 4
Acknowledgements          88                  Supplementary methods: 1
Total                     5858

<div custom-style="Heading, no number">Summary</div>

* We investigated whether global leaf economic relationships across plant functional types (PFTs) are also present within-PFTs, and the extent to which this hierarchical structure can be used to constrain trait estimates.

* We developed a hierarchical multivariate Bayesian model that assumes separate means and covariance structures within and across PFTs and fit this model to seven leaf traits from the TRY database: longevity, specific leaf area, nitrogen content, phosphorus content, dark respiration rate, maximum photosynthetic carboxylation rate ($V_{c,max}$), and maximum photosynthetic electron transport rate ($J_{max}$). 
We evaluated the ability of this model to estimate trait means and reproduce leaf economic trade-offs.

* Trait correlations were generally consistent in direction within and across PFTs, and consistent with predictions of the leaf economic spectrum.
However, correlation strength varied substantially across PFTs indicating that leaf economic relationships within PFTs are often confounded by the unique physiology of certain plant types or environmental conditions in certain biomes.

* Leveraging covariance in multivariate models reduced uncertainties in mean trait estimates, particularly for undersampled trait-PFT combinations.
However, additional constraint from the across-PFT hierarchy was limited compared to the simple multivariate models.

<!--chapter:end:index.Rmd-->

# Introduction

The diversity and dimensionality of the terrestrial biosphere is vast and complex, and therefore there has been a recurring debate in ecology about the utility of reductionist approaches for capturing this variability.
In particular, the use of functional groups with common characteristics has been widely applied in biodiversity studies [@naeem_disentangling_2003] and is essential to the structure of many ecosystem models [@lavorel_plant_1997; @wullschleger_plant_2014].
However, ecologists have long recognized the importance of individual variability and stochasticity in shaping ecosystems
[@gleason_individualistic_1926; @bolnick_why_2011; @rosindell_unified_2011; @clark_why_2016],
and the benefits of more finely-resolved representation of functional diversity for predictive ecology are supported by an increasing body of trait ecology literature
[@mayfield_diversity_2006; @mcmahon_improving_2011; @van_bodegom_going_2012; @reichstein_linking_2014; @violle_emergence_2014; @medlyn_using_2015; @moran_intraspecific_2016].

Plant functional traits can be used to link directly measurable features of individuals to their fitness within an ecosystem and, by extension, ecosystem performance as a whole [@violle_let_2007].
Recent syntheses of global trait databases have revealed that although the functional diversity across plant species is immense, this diversity is constrained by allometries and trade-offs between plant strategies [@wright_worldwide_2004; @kattge_try_2011; @kleyer_why_2015; @diaz_global_2016].
One axis of trait covariation currently receiving attention is the ‘leaf economic spectrum’, which defines a trade-off between plant investment in productive but short-lived leaves versus less productive but sturdy and long-lived leaves [@wright_worldwide_2004; @shipley_fundamental_2006; @reich_world-wide_2014; @diaz_global_2016]
Leaf economic traits are well-correlated with
individual plant productivity [@shipley_functional_2005; @niinemets_within-canopy_2016; @wu_convergence_2016],
litter decomposition rates [@bakker_leaf_2011; @hobbie_plant_2015],
community composition [@burns_patterns_2004; @cavender-bares_multiple_2004],
and ecosystem function [@diaz_plant_2004; @musavi_imprint_2015].
The relative position of plant species along the leaf economic spectrum has been shown to be influenced by climate and soil conditions
[@wright_worldwide_2004; @wright_modulation_2005; @cornwell_community_2009; @ordonez_global_2009; @wigley_leaf_2016].
As a result, relationships between leaf economic traits and climate have been incorporated into ecosystem models to allow for continuous variation in plant function and environmental responses
[@sakschewski_leaf_2015; @verheijen_inclusion_2015].

However, the use of among-trait and trait-environment correlations at the global scale, for both ecological inference and land surface modeling, has several important caveats.
First, observed correlations at the global scale do not always hold at smaller scales (such as sites, species, and individuals).
For example, some studies suggested consistent correlations across scales [@wright_worldwide_2004; @albert_multi-trait_2010; @asner_amazonian_2014]
whereas others showed no or even opposite correlations [@albert_intraspecific_2010; @messier_how_2010; @wright_does_2012; @feng_scale_2013; @grubb_relationships_2015; @wigley_leaf_2016; @messier_traitnetwork; @kichenin_2013_contrasting].
Many mechanisms have been suggested for deviation from global trait relationships at smaller scales.
Trade-offs between strategies may only be applicable when multiple competing strategies co-occur, and in cases where strong environmental filters allow only a narrow range of strategies, alternative processes can drive community assembly [@rosado_trait_dominance; @pierce_csr_cooccur; @grim_pierce_book].
Different selective pressures dominate at different scales, particularly within versus across species [@albert_intraspecific_2010; @messier_how_2010; @kichenin_2013_contrasting],
and the large heterogeneity in the variance structures of traits suggests that different traits have contrasting sensitivity to these different pressures [@messier_trait_2016].
Experimental evidence shows that species can alter different aspects of their leaf economy in a relatively uncoordinated fashion, even when the direction of univariate trait responses to environmental change is consistent [@wright_does_2012].
Meanwhile, across different plant functional types, resource allocation patterns, for instance of nutrients to photosynthesis versus structure and defense, differ substantially, suggesting different investment strategies and varying relationships among traits [@ghimire_2017_Nallocation].
Second, among-trait correlations at any scale do not provide causal evidence for functional trade-offs or even similarity in response to external stimuli [@messier_trait_2016].
Therefore, ascribing too much leverage to trait correlations can lead to an underestimation of plant functional diversity [@grubb_trade-offs_2015].
Third, plants maintain their fitness in a given environment through multiple independent strategies (corresponding to multiple mutually orthogonal axes of trait variability).
As a consequence, changes in key leaf economic traits such as leaf nitrogen content and specific leaf area area may not affect other aspects of plant function, such as
hydraulics [@li_leaf_2015],
overall plant carbon budget [@edwards_leaf_2014],
and dispersal [@westoby_plant_2002].
Finally, modeling ecosystem function based on trait correlations is sampling from the hypothetical space of potential species and communities that could have evolved, rather than constraining models to forecast the actual vegetation we have today as the result of spatial separation and constraints on convergent evolution.
Among other problems, this approach fails to account for the timescales required for adaptation as well as actual limitations of the physiology of different species and community assembly.

An alternative approach is to preserve existing PFT classifications
[though potentially with finer taxonomic, functional, or spatial resolution, e.g. @boulangeat_improving_2012]
while using statistical analyses to account for uncertainty and variability in the aggregated trait values.
For example, the Predictive Ecosystem Analyzer (PEcAn, pecanproject.org), an ecosystem model-data informatics system, parameterizes PFTs using trait probability distributions from a Bayesian meta-analysis of plant trait data across many studies
[@dietze_improving_2013; @lebauer_facilitating_2013].
This approach explicitly separates the processes driving PFT-level differentiation from processes that drive finer-scale functional variability,
and is useful for guiding future data collection and model refinement [@dietze_quantitative_2014].
However, a univariate meta-analysis, like the one currently in PEcAn, is limited by its failure to account for trait correlations, therefore neglecting useful knowledge about relationships across PFTs and between traits.
At the other extreme, existing regional and global scale analyses [e.g. @van_bodegom_going_2012; @sakschewski_leaf_2015] ignore variability within PFTs, often resulting in macroecological, evolutionary, and competitive trade-offs across PFTs being used to drive both acclimation and instantaneous responses within PFTs.

While the leaf economic spectrum has been investigated at the global scale, where it is robust, and at the site or plot scale, where deviations from it are common, it has received less attention at the intermediate scale of PFTs.
Thus, this paper seeks to answer the following questions:
First, to what extent does the leaf economic spectrum hold within vs. across plant functional types?
Second, to what extent can the leaf economic spectrum and other patterns of trait covariance be leveraged to constrain trait estimates, particularly under data limitation?
The answer to these question has implications for both functional ecology and ecosystem modelling.
To evaluate these questions, we develop a hierarchical multivariate Bayesian model that explicitly accounts for across- and within-PFT variability in trait correlations.
We then fit this model to a global database of foliar traits to estimate mean trait values and variance-covariance matrices for PFTs as defined in a major earth system model [Community Land Model, CLM, @clm45_note].
We evaluate the ability of this model to reduce uncertainties in trait estimates and reproduce observed patterns of global trait variation compared to non-hierarchical multivariate and univariate models.
Finally, we assess the generality and scale dependence of trait trade-offs by comparing covariance estimates globally and within each PFT.

<!--chapter:end:01_introduction.md-->

# Materials and Methods

## Trait data

Foliar trait data for this analysis comes from the TRY global traits database (@kattge_try_2011; Tab.\ \@ref(tab:trydata)).
We focused our research on seven foliar traits:
Leaf longevity (months),
specific leaf area (SLA, m$^2$ kg$^{-1}$),
leaf nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
leaf phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
leaf dark respiration at 25°C ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum Rubisco carboxylation rate at 25°C ($V_{c,max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate at 25°C ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
For $V_{c,max}$, we only used values already reported in TRY as being at 25°C.
For $R_{d}$, we normalized the values to 25°C based on reported leaf temperature values following the same methods as @atkin_global_2015.
For $J_{max}$, we normalized the values to 25°C based on reported leaf temperature values using the temperature response function described in @kattge_2007_temperature (Equation 1 therein).
To avoid potential artifacts caused by different trait normalization, we performed analyses separately for both mass- and area-normalized traits [@osnas_global_2013; @lloyd_les].
We restricted our analysis to TRY data that have been quality-controlled, where at least one of the above traits was observed, and for which adequate species information was provided for functional type classification [see @kattge_try_2011].

Although the light- and CO2-saturated photosynthetic rate ($A_{max}$) was an important trait in previous studies, we did not include it in our study for two reasons.
First of all, data on raw photosynthetic rates are highly sensitive to measurement methodology and environmental conditions, which were generally inconsistent or unavailable in TRY.
Second, $A_{max}$ is not a good measure of photosynthetic capacity because it integrates over variability in many physiologically independent traits such as $V_{c,max}$, $J_{max}$, and stomatal conductance, and is therefore not used in vegetation models as a photosynthetic parameter [@Ali_2015].

Following past studies [e.g. @wright_worldwide_2004; @wright_assessing_2005; @onoda_2011_global; @diaz_global_2016], we log-transformed all trait values to correct for their strong right-skewness.

## Plant functional types

We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model [CLM4.5, @clm45_note; Tab.\ \@ref(tab:pfts), Fig.\ 1].
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
Where species attributes disagreed between datasets (e.g., categorized as a shrub in one dataset but a tree in another), we assigned the attribute that was observed most frequently between the datasets (e.g., if five datasets say "shrub" but only one says "tree", we would classify it as a shrub).
Where species lacked data on certain attributes, we assigned those attributes based on higher order phylogeny where appropriate (e.g., *Poaceae* family are usually grasses, *Larix spp.* are deciduous needleleaved trees) or otherwise omitted the species from our analyses.
For biome specification, we collected all latitude and longitude data for each species,
matched these data to 30 second ($\sim1$ km$^{2}$) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
calculated the mean AMT for all sites at which that species was observed,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).


## Multivariate analysis

### Basic model description

In this study, we compared three different models representing different levels of complexity.

The simplest model was the ‘univariate’ model, in which each trait was modeled independently.
For an observation $x_{i,t}$ of trait $t$ and sample $i$:

$$x_{i,t} \sim N(\mu_t, \sigma_t)$$

where $N$ is the univariate normal (Gaussian) distribution with mean $\mu_t$ and standard deviation $\sigma_t$ for trait $t$.

The second-simplest model was the ‘multivariate’ model, in which traits were modeled as samples from a multivariate distribution with a single mean vector and covariance matrix.
For the observed vector of traits ${\mathbf{x_i}}$ for sample $i$:

$$\mathbf{x_i} \sim mvN(\mathbf{\mu}, \mathbf{\Sigma})$$

where $mvN$ is the multivariate normal (Gaussian) distribution with mean vector $\mathbf{\mu}$ and variance-covariance matrix $\mathbf{\Sigma}$.
We ran both of these models independently for each PFT as well as for the entire dataset (as if every observation belonged to the same PFT).

The most complex model was the ‘hierarchical’ model, in which observed trait values were drawn from a PFT-specific multivariate normal distribution describing within-PFT variation and whose parameters were themselves sampled from a global multivariate distribution describing the variation across PFTs.
For the observed vector of traits $\mathbf{x}_{i,p}$ for sample $i$ belonging to PFT $p$:

$$\mathbf{x}_{i,p} \sim mvN(\mathbf{\mu}_p, \mathbf{\Sigma}_p)$$
$$\mathbf{\mu}_p \sim mvN(\mathbf{\mu}_g, \mathbf{\Sigma}_g)$$

where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and variance-covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and variance-covariance matrix describing across-PFT (global) variation.


### Model implementation

We fit the above models using a Gibbs sampling algorithm that leveraged known conjugate prior relationships for efficient exploration of the sampling space.
For priors on all multivariate mean vectors ($\mathbf{\mu}$), we used normal distributions:

$$P(\mathbf{\mu}) \sim mvN(\mathbf{\mu}_0, {\mathbf{\Sigma}}_0)$$

This gives rise to the following expression for the posterior:

$$P(\mathbf{\mu} \mid 
    \mathbf{x}, \mathbf{\Sigma}, 
    \mathbf{\mu}_0, \mathbf{\Sigma}_0)
  \sim
  mvN(\mathbf{\mu^*}, \mathbf{\Sigma^*})$$

$$\mathbf{\Sigma^*} = (\mathbf{\Sigma}_0^{-1} + n \mathbf{\Sigma}^{-1})^{-1}$$
$$\mathbf{\mu^*} = \mathbf{\mu}_0 \mathbf{\Sigma}_0^{-1} + \bar{\mathbf{x}} n \mathbf{\Sigma}^{-1}$$

where ${\bar{{\mathbf{x}}}}$ are the sample means of the data and $n$ is the number of rows in the data.

For priors on all multivariate variance-covariance matrices, we used the Wishart distribution ($W$):

$$P(\mathbf{\Sigma}) \sim W(\nu_0, \mathbf{S}_0)$$

This gives rise to the following expression for the posterior:

$$P(\mathbf{\Sigma} \mid
  \mathbf{x}, \mathbf{\mu},
  \nu_0, \mathbf{\Sigma}_0)
  \sim
  (W(\nu^*, S^*))^{-1}$$

$$\nu^* = 1 + \nu_0 + n + m$$
$$\mathbf{x^*} = \mathbf{x} - \bar{\mu}$$
$$\mathbf{SS} = \mathbf{x^*}^{T} \mathbf{x^*}$$
$$\mathbf{S^*} = (\mathbf{S}_0 + \mathbf{SS})^{-1}$$

where $n$ is the number of rows and $m$ is the number of columns in data matrix $x$. For further details on this derivation, see @gelman_bayesian.

The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating partially missing observations.
Therefore, our algorithm also included multiple imputation of partially missing data, which proceeded as follows:
For a block of data $\mathbf{x'}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
the missing values $\mathbf{x'}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:

$$\mathbf{x'}[m|p] \sim mvN(\mathbf{\mu}', \mathbf{\Sigma}')$$

$$\mathbf{\mu'} = 
  (\mathbf{x'}[p] - \mathbf{\mu'}[p]) 
  (\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])$$
$$\mathbf{\Sigma'} = \mathbf{\Sigma}[m,m] - 
  \mathbf{\Sigma}[m,p]
  (\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])$$
  
We emphasize that the imputation of missing data is performed iteratively as parameters are being estimated, such that each set of imputed values is conditioned on the current sampled state of the parameters.
This approach is therefore distinct from single imputation, where data are imputed first in a separate step prior to parameter estimation.
A demonstration of this multiple imputation approach and how it is used to estimate trait covariance is provided Supporting Information Method S1.

For each model fit, we ran five parallel MCMC chains, continuing the sampling until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1 [@gelman_1992_inference].
We implemented this sampling algorithm in an open source, publicly available R [version 3.4.3; @rstats] package (<http://github.com/ashiklom/mvtraits>).


### Analysis of results

To assess the impact of multivariate and hierarchical constraint on trait estimates,
we compared the mean and 95% confidence intervals of trait estimates for each PFT from each model (Fig.\ 2, Tab.\ S1 and S2).
For reference, we also added the default parameter values of CLM 4.5 [Table 8.1 in @clm45_note] for SLA, $N_{mass}$, $N_{area}$, $V_{c,max,mass}$, and $V_{c,max,area}$ to Fig.\ 2.
To convert CLM's reported C:N ratio to $N_{mass}$, we assumed a uniform leaf C fraction of 0.46.
We then divided this calculated $N_{mass}$ by the reported SLA to obtain $N_{area}$.
We calculated $V_{c,max,mass}$ by multiplying the reported $V_{c,max,area}$ by the reported SLA.

To test the hypothesis that the multivariate and hierarchical models offer more value in terms of uncertainty constraint at smaller sample sizes, we calculated the relative uncertainty ($\alpha$) as a function of the mean ($\mu$) and upper ($q_{0.975}$) and lower ($q_{0.025}$) confidence limits of trait estimates.

$$ \alpha = \frac{q_{0.975} - q_{0.025}}{\mu} $$

We then fit a generalized linear model relating relative uncertainty to sample size ($n$) for each of the model types (univariate, multivariate, and hierarchical; Fig. 3).

$$ \log{\alpha} = b_0 + b_1 \log{n} $$

If all three models performed equally well at all sample sizes, their respective slope and intercept coefficients would be statistically indistinguishable.
On the other hand, models that perform better should have
lower intercept ($b_0$) coefficients, indicating generally lower uncertainty,
and
lower slope ($b_1$) coefficients, indicating a reduced sensitivity of uncertainty ($\alpha$) to sample size ($n$).

To assess the consistency of within- and across-PFT trait trade-offs, we looked at covariance estimates for each trait pair and, where these values were significantly different from zero ($p < 0.05$),
we calculated the eigenvalues from the pairwise variance-covariance matrix for that trait pair and plotted the corresponding dominant eigenvectors centered on the mean estimates (Fig.\ 4).
This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance, @cornwell_community_2009 and @albert_intraspecific_2010.
Due to the small number of points used to estimate across-PFT covariance in the hierarchical model, none of the across-PFT covariances estimated in the hierarchical model were significantly different from zero ($p < 0.05$).
For this reason, we compared within-PFT covariances as estimated by the hierarchical model with the covariances estimated by fitting a multivariate model to all of the data.

Besides the consistency in the direction of trait covariance globally and between different PFTs, we also investigated the strength and predictive power of these covariances, which is represented by correlation coefficients (i.e. the pairwise covariance normalized to the variances of the component variables).
To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.\ 5).

The magnitude of correlation coefficients can be strongly influenced by data sampling, particularly sample size and range (correlations are likely to be stronger when data have more samples and a larger range).
To evaluate the contribution of uneven data sampling to our correlation predictions, we plotted the square of each pairwise correlation coefficient against pairwise sample size and normalized data range (Fig.\ S1 and S2).

The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework at <https://osf.io/w8y73/>.
The TRY data request used for this analysis has been archived at <http://try-db.org>,
and can be retrieved by providing the TRY data request ID (#1584).
Alternatively, the exact pre-formatted data used in this analysis are available on request to Alexey Shiklomanov (ashiklom@bu.edu).

<!--chapter:end:02_methods.Rmd-->

# Results

## Estimates of PFT-level means

In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig.\ 2, Tab.\ S1 and S2).
Where we observed differences between models, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
Significant differences in trait estimates between univariate and multivariate models occurred even for traits with relatively large sample sizes, such as leaf nitrogen content.

Evergreen PFTs had by far the largest leaf lifespan, with the longest lifespan observed for temperate and boreal needleleaf evergreen species.
Meanwhile, all of deciduous species had lifespans shorter than 7 months.
Among deciduous species, lifespan was generally longer in warmer biomes than colder ones.

Across-PFT patterns in SLA and $N_{mass}$, $P_{mass}$, and $R_{d,mass}$ were similar.
Temperate broadleaved deciduous trees and shrubs generally had among the highest values of these traits, while temperate evergreen trees and shrubs had generally among the lowest.
However, none of these patterns were universal to all four traits.
For example, tropical evergreen trees had relatively high $N_{mass}$ and mean SLA and $R_{d,mass}$, but among the lowest $P_{mass}$.
Similarly, temperate and boreal shrubs had higher $N_{mass}$ and $P_{mass}$ than any of the grasses, but comparable SLA.

Across-PFT patterns in $N_{area}$, $P_{area}$, and $R_{d,area}$ were different from their mass-normalized counterparts.
For example, tropical broadleaved evergreen and needleleaf evergreen trees had among the lowest $N_{mass}$ and $P_{mass}$ basis but among the highest $N_{area}$ and $P_{area}$, while the opposite was true of deciduous temperate trees and shrubs.
Species with N contents near the middle of the observed range did not shift as dramatically depending on type of normalization.

C3 grasses had both the highest $V_{c,max,mass}$ and $V_{c,max,area}$.
Compared to broadleaved trees, temperate needleleaved evergreen trees had lower $V_{c,max,mass}$ but higher $V_{c,max,area}$.
Among broadleaved trees, deciduous trees had higher $V_{c,max,mass}$ and slightly higher $V_{c,max,area}$ than evergreen trees.
Between the deciduous and evergreen tree PFTs, we observed no significant trend by climate zone.

C3 grasses and temperate needleleaved evergreen trees had the highest $J_{max,area}$, but temperate broadleaved deciduous trees had the highest $J_{max,mass}$.
All of the shrub PFTs had the lowest $J_{max,mass}$ but average or above-average $J_{max,area}$, while the opposite was true of broadleaved tropical PFTs.
Of the tree PFTs, needleleaved evergreen trees had the highest $J_{max,area}$ but the lowest $J_{max,mass}$.

A key application of this study was to provide data-driven parameter estimates for Earth System Models.
To this end, we compared our mean parameter estimates with corresponding default parameters in CLM 4.5 [@clm45_note] (Fig.\ 2).
Our mean estimates of SLA agreed with CLM's defaults [@clm45_note, Table 8.1] only for tropical broadleaved evergreen trees, and for all other PFTs, our estimates are significantly lower.
For $N_{mass}$, our estimates agreed reasonably well with CLM for evergreen temperate trees, needleleaved trees, and C3 arctic grasses, and were substantially different for all other PFTs.
Our $N_{mass}$ estimates also varied much more across PFTs than CLM's parameters.
For $N_{area}$, our estimates were significantly higher than CLM's for all PFTs, likely due to CLM's overestimates of SLA.
Our estimates of $V_{c,max_mass}$ were lower across all PFTs, with particularly large differences for tropical and temeprate broadleaf deciduous trees and evergreen shrubs, and temperate C3 grasses.
Our estimates of $V_{c,max,area}$ showed better agreement, though our values were still significantly lower for many PFTs.
Like us, @kattge_2009_vcmax also found that $V_{c,max,area}$ was overestimated by Earth System models,
but their estimates of $V_{c,max,area}$ and $N_{area}$ are generally slightly higher than ours.

We observed clear differences in the relative uncertainties of mean estimates with respect to sample size.
All of the high-latitude PFTs consistently had among the largest error bars around their mean estimates relative to other PFTs, while the traits with the largest uncertainties were dark respiration, $V_{c,max}$, and $J_{max}$.
For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model both reduced uncertainty overall (lower intercept) and reduced the sensitivity of estimate uncertainty to sample size (lower slope) compared to the univariate model (Fig.\ 3).
However, this analysis revealed no consistent significant benefit from the hierarchical model.


## Trait correlation patterns across- and within-PFTs

For all traits except leaf lifespan, pairwise trait correlations were generally consistent in direction both globally and within each PFT (Fig.\ 4).
In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
The same was generally true of correlations of mass-based traits with leaf lifespan, but correlations of leaf lifespan with area-normalized traits were more variable.
The correlation between $N_{area}$ and leaf lifespan was positive globally and for evergreen shrubs, tropical broadleaved deciduous trees, temperate needleleaved evergreen trees but negative for temperate and boreal broadleaved deciduous trees and not significant for any other PFTs.
Similarly, the correlation between $P_{area}$ and leaf lifespan was positive globally but negative for evergreen shrubs and not significant for any other PFTs.
The correlation between leaf lifespan and $R_{d,area}$ was significant and negative globally, but was not significant within any PFTs.
The only significant correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees.

The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ 5).
In some cases, this variability was driven by low sample sizes (Fig.\ 1, S1; Tab.\ S3, S4).
For instance, needleleaved deciduous trees, the most undersampled PFT in our analysis, were often the only PFT for which a correlation was not statistically significant.
Similarly, we had no observations of dark respirations for deciduous boreal shrubs, which explains why we found no significant correlations of dark respiration with any other trait for that PFT.
That being said, the relationship between correlations strength and sample size was very inconsistent (Fig.\ S1; Tab.\ S4).
For example, temperate needleleaf evergreen trees and C3 arctic grasses showed the strongest $N_{area}$-$P_{area}$ correlations despite having relatively average sample sizes for both traits.
In fact, every trait pair we investigated had at least one case (and often several cases) where a PFT with a larger sample size showed weaker correlations than PFTs with lower sample sizes, or where correlation strength varied significantly among PFTs with similar sample sizes (Fig.\ S1).
Relationships between correlation strength and data range were even less consistent (Fig.\ S2).
Based on these results, we conclude that the variation we observed in our correlation results is not driven primarily by sampling artifacts and at least partially captures some underlying ecophysiological differences between PFTs.

<!--chapter:end:03_results.Rmd-->

# Discussion

## Scale dependence of the leaf economic spectrum

The canonical leaf economic spectrum is defined by a negative correlation of between SLA and leaf lifespan, and a positive correlation of SLA with $N_{mass}$, $P_{mass}$, and photosynthesis and respiration rates [@wright_worldwide_2004].
The first objective of this paper was to investigate the extent to which these relationships hold within and across PFTs.
Our results indicate that the leaf economic spectrum generally still holds within PFTs, at least at the functional and phylogenetic scale of the current generation of Earth System Models.
Within PFTs, correlations between SLA, $N_{mass}$, and $P_{mass}$ were consistently positive, and correlations of these traits with leaf lifespan were generally negative (though, for many PFTs, correlations were not significantly different from zero).
Although we did not include maximum photosynthesis rate ($A_{max}$), $V_{c,max,mass}$ and $J_{max,mass}$ generally exhibited the expected positive correlations with SLA and negative correlations with leaf lifespan, as did $R_{d,mass}$, though many correlations were not significant, primarily due to sample size.

While the direction of relationships we observed was largely consistent across different PFTs, the strength of these correlations was more variable.
For example, the correlations of SLA with $N_{mass}$ and $P_{mass}$ were generally weaker in needleleaved PFTs than in broadleaved PFTs.
On the other hand, relationships of SLA and $N_{area}$ were strongly negative for all PFTs (except the data-limited needleleaved deciduous trees), and temperate needleleaved species in particular had among the strongest SLA-$N_{area}$ and SLA-$P_{area}$ correlations of any PFT.
Taken together with the finding that evergreen conifers are relatively unique in their consistently proportional allocation of N to cell walls and Rubisco [@onoda_physiological_2017], our results support the idea that needleleaved species primarily adapt to environmental changes through structural changes (i.e. increasing or decreasing SLA) rather than through modulating foliar biochemistry [@robakowski_2004_needle].

We also found that the strength of correlations between leaf nutrient concentrations and traits related to photosynthetic metabolism ($V_{c,max}$ and $J_{max}$)---which are often used in ecosystem modeling to parameterize photosynthesis [@clm45_note; @rogers_roadmap_2017]---were highly variable between PFTs.
Although trait correlations are not necessarily indicative of allocation strategies, this result generally supports the findings of @ghimire_2017_Nallocation that N allocation strategies to photosynthesis vary widely by plant functional type.
In particular, the correlations of $N_{mass}$ with traits related to photosynthesis and respiration ($R_{d,mass}$, $V_{c,max,mass}$, and $J_{max,mass}$) were weaker for tropical evergreen broadleaved trees than for most other PFTs, but the correlation of these traits with $P_{mass}$ was comparable or even slightly higher.
This suggests that the productivity of tropical species is generally P-limited [@reich_global_2004; @ghimire_2017_Nallocation], that N allocation strategies are more variable under N-poor than N-rich conditions [@ghimire_2017_Nallocation], or more generally that photosynthetic parameters are affected more strongly by environmental covariates than leaf nitrogen contents [@Ali_2015].
Meanwhile, our result that the $N_{area}$ - $V_{c,max,area}$ correlation is generally weaker for needleleaved than broadleaved species was also found by @kattge_2009_vcmax, and supports the lower allocation of N to photosynthesis found by @ghimire_2017_Nallocation.
In light of this fact and considering that boreal forests, largely populated by needleleaf species, have the largest influence on global climate of any biome [@Snyder_2004; @bonan_forests_2008], we suggest that parameterization of needleleaf tree productivity based on foliar nitrogen content in Earth System Models be treated with caution.

Correlations of all traits with leaf lifespan were weaker (often to the point of becoming insignificant) within most PFTs than globally.
This suggests that leaf economic relationships related to leaf lifespan are dominated by the fundamental differences between deciduous and evergreen PFTs, while the factors driving variability in leaf lifespan within PFTs are more complex and idiosyncratic [@Reich_2014; @wu_leaf_2016].
That being said, much of this within-PFT variability is known to be driven by variations in shade responses, and a key limitation of our study is that we have no information about the relative canopy positions at which traits were collected [@lusk_why_2008; @keenan_global_2016].

Across PFTs, the interaction between growth form and biome in PFT definitions (Table 1) confounds the interpretation of our results with respect to well established ecological and biogeographic patterns.
For example, @poorter_causes_2009 found variation in SLA by growth form was, from highest to lowest, grasses, deciduous woody plants, evergreen woody plants, and succulents, while the pattern by biome was, again from highest to lowest, grassland, tundra, tropical forest, temperate forest, woodland, shrubland, and desert.
We observed as expected that arctic grasses had lower mean SLA than temperate grasses, and that evergreen trees had lower SLA than their deciduous counterparts.
However, we observed by far the highest mean SLA values in temperate deciduous broadleaf trees, rather than in grass PFTs as expected.
Similarly to @onoda_2011_global, we found no consistent patterns in SLA with temperature:
Among broadleaved evergreen PFTs, temperate species had lower SLA than tropical, but among broadleaved deciduous PFTs, temperate species had higher SLA than both tropical and boreal species.
With respect to the expected decline in foliar N:P ratios with latitude [@reich_global_2004], we found that $N_{mass}$ was generally higher in PFTs associated with colder biomes compared to warmer ones, but observed biome-related differences in $P_{mass}$ only among deciduous broadleaved and needleleaved evergreen PFTs.
Contrary to @atkin_global_2015, our results for both $R_{d,mass}$ and $R_{d,area}$ failed to show a trend with respect to biome.
However, this comparison may not be entirely fair because our study design inherently averages over the extensive climatic variability within PFTs.

Finally, there has been some debate in the trait ecology community about the use of mass- or area-normalized traits in analyses of the leaf economic spectrum.
Two studies [@osnas_global_2013; @lloyd_les] independently concluded that leaf economic relationships among mass-based traits emerge inevitably out of variation in SLA and are therefore not ecologically meaningful.
Responses to these criticisms have suggested that both mass- and area-based normalization have merit, as mass-based traits have a natural interpretation in terms of investment costs and allocation while area-based traits are important due to the fundamentally area-based nature of energy and gas fluxes through leaf surfaces [@westoby_lloyd_response; @poorter_les_response].
Our study suggests the latter, that investigation of trait correlations on both a mass- and area-basis can yield biologically and ecologically meaningful conclusions.
For one, our discussion of differences in ecological strategies between broadleaved and needleaved species fundamentally depends on comparative analysis of mass- and area-normalized nutrient contents.
Meanwhile, our discussion of tropical tree productivity with respect to foliar nutrient contents is generally supported by both mass- and area-normalized traits (i.e. the correlation between $P$ and $V_{c,max}$ is stronger than the correlation of $N$ and $V_{c,max}$ for tropical species whether we use the mass- or area-normalized versions of these traits).


## Covariance as constraint

The second objective of this paper was to investigate the ability of trait covariance to provide additional information that could be used to reduce uncertainties in trait estimates.
In agreement with our expectations, our results show that accounting for trait covariance constrained the uncertainty around PFT-level trait means, particularly for trait-PFT combinations with low sample sizes (Fig.\ 2 and 3).
Moreover, accounting for trait covariance occasionally also changed the position of trait mean estimates, even for PFT-trait combinations with relatively large sample sizes (e.g. $N_{mass}$ for temperate broadleaved deciduous trees, Fig.\ 2).
This result echoes @diaz_global_2016 in demonstrating the importance of studying the multivariate trait space rather than individual traits. 
Such shifts suggest that the sampling of these traits in the TRY database is not representative, which is evident from looking at the relative sample sizes of different traits (Fig.\ 1; see also Kattge et al. -@kattge_try_2011). 
These shifts also indicate that parameter estimates based on univariate trait data [e.g. @lebauer_facilitating_2013; @dietze_quantitative_2014; @butler_2017_mapping] may not only be overestimating uncertainty, but may also be systematically biased.
Although our results clearly show that many time- and labor-intensive traits, such as $R_{d}$, $V_{c,max}$, and $J_{max}$, still lack the observations to estimate covariance with other traits for certain PFTs,
our results also show that the effective sample size of all traits is enhanced when covariance is taken into account.
For example, field and remote sensing studies that estimate only certain traits like SLA and $N_{mass}$ but not others (such as $P_{mass}$ and $R_{d,mass}$) can leverage trait correlations to provide indirect constraint [@serbin_spectroscopic_2014; @singh_imaging_2015; @musavi_imprint_2015; @lepine_examining_2016].
We therefore suggest that future observational campaigns consider trait covariance when deciding which traits to measure.

The additional benefit of hierarchical multivariate modeling in our study was limited, largely due to the relatively low number of points used to estimate that across-PFT covariance structure.
Therefore, for parameterizing the current generation of ecosystem models using well-sampled traits, we suggest that simple multivariate models fit independently to each PFT are sufficient and that the additional conceptual challenges and computational overhead of hierarchical modeling are not required.
However, for modeling work that requires larger numbers of PFTs [@boulangeat_improving_2012], and especially for models that are parameterized on the basis of individual species [e.g. Linkages, @linkages], the benefits of hierarchical modeling may accumulate [@clark_why_2005; @dietze_capturing_2008; @cressie_accounting_2009; @webb_structured_2010].

More generally, we foresee tremendous potential of multivariate and hierarchical modeling approaches to elucidating the role of functional traits in organismal and ecosystem function. 
Besides the many important foliar traits that we did not include (e.g.\ pigment contents, leaf hydraulic traits), our approach could readily be applied to other plant traits whose relationship to the leaf economic spectrum is less clear. 
One example is hydraulic traits:
While stem and leaf hydraulic traits are correlated [@bartlett_hydraulic], a potential scaling between hydraulic and leaf economic traits is poorly understood [@reich_world-wide_2014; @li_leaf_2015]. 
Similarly, reexamining the relationships defining wood [@chave_woodeconomics; @baraloto_decoupled_2010; @fortunel_leaf_2012] and root [@kramer-walter_root_2016; @valverde-barrantes_root_2016] economic spectra, as well as their relationship to the foliar traits, would provide useful information on possible scale-dependence of plant growth and allocation strategies.
We emphasize that the relative difficulty of measuring hydraulic and other non-foliar traits [e.g @jansen_hydraulic_traits] further increases the value of any technique that can fully leverage the information they provide. 
More generally, multivariate and hierarchical modeling has the potential to reveal functional trade-offs that are mutually confounding at different scales, thereby enhancing our understanding of processes driving functional diversity.

## Conclusions

The tremendous functional diversity of plants continues to be a major challenge for functional ecology and ecosystem modeling.
Functional diversity research fundamentally depends on dimensionality reduction through a search for meaningful pattern that can be exploited to take reasonable guesses at average behavior. 
The trait trade-offs comprising the leaf economic spectrum are one such pattern. 
In this paper, we reaffirm the existence of the leaf economic spectrum both globally and, with some caveats, within plant functional types typically used in the current generation of Earth System Models.
We also highlight how the strength of leaf economic relationships can be influenced by biotic and abiotic factors specific to certain PFTs.
Finally, we show how patterns of trait covariance like the leaf economic spectrum can be leveraged to inform trait estimates, particularly at small sample sizes.

<div custom-style="Heading, no number">Acknowledgements</div>

This project was supported by NASA grant NNX14AH65G and NSF grants 1261582 and 1458021, as well as the TRY initiative on plant traits (http://www.try-db.org).
The TRY initiative and database is hosted, developed and maintained by J. Kattge and G. Boenisch (Max Planck Institute for Biogeochemistry, Jena, Germany).
TRY is currently supported by DIVERSITAS/Future Earth and the German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig.
The authors would also like to thank Christine Rollinson, Istem Fer, and Colin Averill for their valuable feedback on early drafts of this manuscript.

<div custom-style="Heading, no number">Author contributions</div>

ANS wrote the manuscript and implemented the analysis.
ANS and EMC designed the analysis and figures.
MCD conceived the original idea for the manuscript, guided its development, and provided financial support.
MB, SJ, KK, ÜN, and NAS provided extensive feedback on multiple drafts of the manuscript, and contributed data.
All other authors contributed data.

<!--chapter:end:04_discussion.md-->

<div custom-style="Heading, no number">Tables</div>

```{r "knitr_options", echo = FALSE, message = FALSE, warning = FALSE, error = FALSE}
library(knitr)
library(here)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE, error = FALSE
)
options(knitr.table.format = "pandoc")
```

<!-- Tables ---------------------------->
```{r "read_chunks"}
knitr::read_chunk(here::here("scripts/manuscript-analyses/try_ref_table.R"))
knitr::read_chunk(here::here("scripts/manuscript-analyses/table_n_species.R"))
```

<!-- TRY data citations -->
```{r "trydata"}
```

<!-- Number of species in each PFT -->
```{r "pfts"}
```

<!--chapter:end:05_figures_tables.Rmd-->

